A free/open-source machine translation system for English to Kazakh

### 2015-03-06 engkaz-eamt-2015/engkaz-eamt-2015.tex
foo
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59241 72bbbca6-d526-0410-a7d9-f06f51895060

\author{Aida Sundetova\\
  Kazakh National University,  \\
  050040 Almaty, \\
  {\tt xxx@xxx}\\
  \And
  \textbf{Francis M. Tyers}\\
  HSL-fakultetet\\ 
  UiT Norgga \'{a}rktala\v{s} universitehta \\
  9017 Romsa, \\
  {\tt francis.tyers@uit.no}
  \And
  Mikel L. Forcada\\
  Dept. Lleng. i Sist. Inform. \\
  Universitat d'Alacant, \\ 
  Alacant \\
  {\tt mlf@dlsi.ua.es}  
}

@@

Current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
contains 13,135 entries, \todo{13130 before!-changed} There are 168 ``chunker'' rules and 148 interchunk rules. 
Evaluating was tested against revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 
over EuroParl,\footnote{\url{http://www.statmt.org/europarl/v7/es-en.tgz}} SETimes,\footnote{\url{http://nlp.ffzg.hr/data/corpora/setimes/setimes.en-tr.txt.tgz}} NewsCommentary,\footnote{\url{http://www.statmt.org/wmt13/training-parallel-nc-v8.tgz}} Wikipedia.\footnote{\url{http://dumps.wikimedia.org/enwiki/20140402/enwiki-20140402-pages-articles.xml.bz2}; last checked 04/06/2014} 
Table~\ref{table:coverage} presents the size of the corpora and the vocabulary coverage of the system for the given corpus.

@@

As can be seen from Table~\ref{table:coverage}, the coverage of dictionaries is good, average score is 94\%. Our system outperforms the other 
available RBMT system, but fall short of the state-of-the-art performance.\todo{represented by...} However, unlike the state of 
the art, which uses unavailable corpora, our system is free and open and can be improved.

>>>>

\author{Aida Sundetova\\
  Kazakh National University,  \\
  050040 Almaty, \\
  {\tt sun27aida@gmail.com}\\
  \And
  \textbf{Francis M. Tyers}\\
  HSL-fakultetet\\ 
  UiT Norgga \'{a}rktala\v{s} universitehta \\
  9017 Romsa, \\
  {\tt francis.tyers@uit.no}
  \And
  Mikel L. Forcada\\
  Dept. Lleng. i Sist. Inform. \\
  Universitat d'Alacant, \\ 
  Alacant \\
  {\tt mlf@dlsi.ua.es}  
}

@@

Current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
contains 13,135 entries. There are 168 ``chunker'' rules and 148 interchunk rules. 
Evaluating was tested against revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 
over EuroParl,\footnote{\url{http://www.statmt.org/europarl/v7/es-en.tgz}} SETimes,\footnote{\url{http://nlp.ffzg.hr/data/corpora/setimes/setimes.en-tr.txt.tgz}} NewsCommentary,\footnote{\url{http://www.statmt.org/wmt13/training-parallel-nc-v8.tgz}} Wikipedia.\footnote{\url{http://dumps.wikimedia.org/enwiki/20140402/enwiki-20140402-pages-articles.xml.bz2}; last checked 04/06/2014} 
Table~\ref{table:coverage} presents the size of the corpora and the vocabulary coverage of the system for the given corpus.

@@

As can be seen from Table~\ref{table:coverage}, the coverage of dictionaries is good, average score is 94\%. Our system outperforms the other 
available RBMT system, but fall short of the state-of-the-art performance represented by Google Translate. However, unlike the state of 
the art, which uses unavailable corpora, our system is free and open and can be improved.
